---
title: "P8105 Homework 6 - Linear Models"
author: "Sydney Ng"
date: "due December 9, 2020 at 10:00 PM ET"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(p8105.datasets)
library(purrr)
library(mgcv)

knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = .6,
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

```{r message=FALSE, warning=FALSE}
homicide_df <-
  read_csv("./homicide-data.csv") %>%
  mutate(
    victim_age = as.numeric(victim_age),
    city_state = str_c(city, state, sep = "_"),
    resolved = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1
    )
  ) %>%
  filter(city_state != "Tulsa_AL",
         victim_race %in% c("Black", "White")) %>%
  select(city_state, resolved, victim_age, victim_race, victim_sex)

# by default, R will remove missing values (the whole row) when we want
# to run a regression model, unless we make it some unknown category...
```

Starting with one city
```{r}
baltimore_df <-
  homicide_df %>%
  filter(city_state == "Baltimore_MD")

glm(resolved ~ victim_age + victim_race + victim_sex,
    data = baltimore_df,
    family = binomial()) %>%
  broom::tidy() %>%
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96*std.error),
    CI_upper = exp(estimate + 1.96*std.error)
  ) %>%
  select(term, OR, starts_with("CI")) %>%
  knitr::kable(digits = 3)
```

Try this across different cities
```{r}
model_results_df <-
  homicide_df %>% 
  nest(data = -city_state) %>% # 3 cities do not report race and won't be included
  mutate(
    models = map(.x = data, 
                 ~glm(resolved ~ victim_age + victim_race + victim_sex,
                     data = .x,
                     family = binomial())),
    results = map(models, broom::tidy)
  ) %>%
  select(city_state, results) %>%
  unnest(results) %>%
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96*std.error),
    CI_upper = exp(estimate + 1.96*std.error)
  ) %>%
  select(term, OR, starts_with("CI"))
```

```{r}
model_results_df %>%
  filter(term == "victim_sexMale") %>%
  mutate(city_state = fct_reorder(city_state, OR)) %>%
  ggplot(aes(x=city_state, y=OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


## Problem 2

In my linear regression model, we are using 

* `fincome` and `mrace` because SES and race are common confounders in epidemiological studies. Since mother's race is categorical, dummy variables are created in the regression and "White" is treated as the reference category.
* `catbmi` because previous studies show that underweight and obese mothers significantly influence baby birthweight; I made `ppbmi` into a categorical variable using the standard cut-offs for underweight and obese categories. Normal/overweight is treated as a single reference category.
* `gaweeks` because how far along into the pregnancy (e.g., premature births) can influence baby birthweight.
* `malform` because it is defined as the presence of malformations that could affect weight.
* and `momage` because higher age is associated with risk of lower birthweight.

I decided not to include variables `ppwt` and `mheight` because both weight and height get factored into BMI. I also debated including `parity` since nulliparous women are more likely to have lower birthweight babies, but also decided against this because it could be correlated with age (i.e., the older the woman is, the more likely she will not be nulliparous especially if she appears in this specific dataset).

```{r message=FALSE}
birthweight_df <-
  read_csv("./birthweight.csv") %>%
  mutate(
    babysex = if_else(babysex == 1, "Male", "Female"), # no missing values
    mrace = case_when(mrace == 1 ~ "White", mrace == 2 ~ "Black",
                      mrace == 3 ~ "Asian", mrace == 4 ~ "Puerto Rican",
                      mrace == 8 ~ "Other"),
    frace = case_when(frace == 1 ~ "White", frace == 2 ~ "Black",
                      frace == 3 ~ "Asian", frace == 4 ~ "Puerto Rican",
                      frace == 8 ~ "Other", frace == 9 ~ "Unknown"),
    malform = if_else(malform == 1, "Present", "Absent"),
    mrace = as.factor(mrace),
    mrace = relevel(mrace, ref = "White"),
    catbmi = case_when(ppbmi < 18.5 ~ "Underweight",
                       ppbmi > 30 ~ "Obese",
                       ppbmi <= 30 & ppbmi >= 18.5 ~ "Normal/Overweight"),
    catbmi = as.factor(catbmi),
    catbmi = relevel(catbmi, ref = "Normal/Overweight")
  ) %>%
  select(-ppbmi)

my_model <- glm(bwt ~ 
                  fincome + mrace + catbmi + momage + gaweeks + malform + smoken,
    data = birthweight_df,
    family = gaussian) 

my_model %>%
  broom::tidy() %>%
  mutate(
    CI_lower = estimate - 1.96*std.error,
    CI_upper = estimate + 1.96*std.error
  ) %>%
  select(term, estimate, starts_with("CI")) %>%
  knitr::kable(digits = 3)
```

Adding predictions and residuals

```{r}
my_pred <- add_predictions(birthweight_df, my_model) %>%
  add_residuals(my_model) %>%
  select(fincome, mrace, catbmi, momage, gaweeks, malform, smoken, pred, resid)

my_pred %>%
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Residuals vs Predicted Values",
    x = "Predicted Birthweight (grams)",
    y = "Residual",
    caption = "A plot of model residuals against fitted values of baby birthweights including \nfamily income, mother race, BMI, age, gestational age, malformatons, and smoking.")
```


#### Model 2: Model including length at birth and gestational age
```{r}
model2 <- glm(bwt ~ blength + gaweeks,
    data = birthweight_df,
    family = gaussian) 

model2 %>%
  broom::tidy() %>%
  mutate(
    CI_lower = estimate - 1.96*std.error,
    CI_upper = estimate + 1.96*std.error
  ) %>%
  select(term, estimate, starts_with("CI")) %>%
  knitr::kable(digits = 3)
```


#### Model 3: Modeling including head circumference, length, sex, and all interactions (including the three-way interaction) 

```{r}
model3 <- glm(bwt ~ bhead*blength*babysex,
    data = birthweight_df,
    family = gaussian) 

model3 %>%
  broom::tidy() %>%
  mutate(
    CI_lower = estimate - 1.96*std.error,
    CI_upper = estimate + 1.96*std.error
  ) %>%
  select(term, estimate, starts_with("CI")) %>%
  knitr::kable(digits = 3)
```

#### Comparing models using Cross-Validation
```{r message = FALSE}
# splitting training and testing data by 80%/20%
train_df = sample_frac(birthweight_df, .80)
test_df = anti_join(birthweight_df, train_df)

crossval_df <- crossv_mc(birthweight_df, 100) %>%
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

crossval_df <-
  crossval_df %>%
    mutate(
      my_mod = map(train, 
                   ~glm(bwt ~ fincome + mrace + catbmi + momage + gaweeks +
                          malform + smoken, family = gaussian, data = .x)),
      mod_2 = map(train, ~glm(bwt ~ blength + gaweeks, family = gaussian, 
                              data = .x)),
      mod_3 = map(train, ~glm(bwt ~ bhead*blength*babysex, family = gaussian,
                              data = .x)))%>% 
    mutate(
      rmse_my_mod = map2_dbl(my_mod, test, ~rmse(model = .x, data = .y)),
      rmse_mod_2 = map2_dbl(mod_2, test, ~rmse(model = .x, data = .y)),
      rmse_mod_3 = map2_dbl(mod_3, test, ~rmse(model = .x, data = .y))) %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model))

crossval_df %>% 
  ggplot(aes(x = model, y = rmse)) +
  geom_violin(aes(fill = model, alpha = 0.5)) +
  theme(legend.position = "none") +
  labs(
    title = "Violin Plot Comparing RMSE of Models Predicting Baby Birthweight",
    x = "Model",
    y = "RMSE",
    caption = "Comparing my model, Model 2 including length at birth and gestational age, and Model \n3 including head circumference, length, sex, and all interactions using cross-validation")
```

Using the root-mean squared error (RSME) to compare the three models, it turns out that my model is actually pretty bad at predicting baby birthweights using the all the variables I included. Model 3 including baby head circumference, length, sex, and all interactions is the best model for prediction, whereas Model 2 including length at birth and gestational age is a close second.

## More about problem 3

```{r message=FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

A simple linear regression with `tmax` as the outcome and `tmin` as the predictor.

```{r message=FALSE}
bootstrap_results <-
  weather_df %>%
  bootstrap(n = 50) %>%
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x)),
    results = map(models, broom::tidy),
    diagnostics = map(models, broom::glance)
  ) %>%
  unnest(diagnostics) %>%
  select(.id, results, r.squared) %>%
  unnest(results)

results_df <-
  bootstrap_results %>%
  group_by(.id, r.squared) %>%
  summarize(
    log_b0b1 = log(sum(estimate))
  )

CI_df <-
  bootstrap_results %>%
  group_by(.id) %>%
  summarize(CI_lower = quantile(estimate, 0.025),
            CI_upper = quantile(estimate, 0.975))


```


```{r}
# central park weather dataset, getting rid of Waiki and Wtaerhole
# max temp and min temp scatterplot, we are interested in the r^2

# log(beta0*beta1)
# fit a model using the data and compute the adjusted r^2 using broom::glance()
# then use broom::tidy() to get all the coefficients to be able to multiply them
# AFTER you get the process, use bootstrapping
```


